Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-01 paper/paper_new.tex

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source particularly for researchers seeking use [apply] machine learning to the vast collections of digitized specimens. Unfortunately, the metadata available is often sparse and, at times, erroneous; and it is prohibitively expensive to address these limitations through traditional, manual [*human] means. This paper reports on a research applying machine-driven [*automatic] approaches to analyze digitized fish images and extract a variety of important features. The digitized fish specimen are being analyzed as part of the XXX project [*Biology Guided Neural Networks (BGNN)], where researchers are developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. The paper reports on research combining machine  learning  and  image  informatics  tools and techniques. A sample 7247 images from the INHS (write-out) were examined via automatic processes; and results show we can  accurately determine metadata, such as, fish quantity, fish orientation, image scaling based on ruler identification and measurement, and general image quality metrics. Results also indicate that automatic approaches "can outperform humans" [and/or "expedite"] the creation of metadata that's critical to leveraging digitized specimens for machine learning. The sections that follow cover background context, reviews the research design, reports results Next,  the  results  are presented, followed by a discussion. 
%items in square brackets are just other suggested words, e.g., select one or the other, or .... 
%something about implications above
%something about comparison to human results.

@@

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms \cite{dc_terms}, the Ecological Metadata Language (EML) \cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.

This limitation is magnified when trying get a handle on the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community advocated for data fitness standards \cite{chapman2020developing}. Even so, metadata properties used to record image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, justifying the capture of 22 metadata properties supporting selection of high quality images for the development of a bgnn. Although, the manual generation of this metadata is costly in terms of labor and time.

>>>>

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source particularly for researchers seeking use [apply] machine learning to the vast collections of digitized specimens. Unfortunately, the metadata available is often sparse and, at times, erroneous; and it is prohibitively expensive to address these limitations through traditional, manual [*human] means. This paper reports on a research applying machine-driven [*automatic] approaches to analyze digitized fish images and extract a variety of important features. The digitized fish specimen are being analyzed as part of Biology Guided Neural Networks (BGNN) initiative, where researchers are developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. The paper reports on research combining machine  learning  and  image  informatics tools and techniques. A sample 7247 images from the Illionois Natural History Survey (INHS) were examined via automatic processes [*<--needs help]. Results show we can  accurately determine metadata, such as, fish quantity, fish orientation, image scaling based on ruler identification and measurement, and general image quality metrics. Results also indicate that automatic approaches "can outperform humans" [*and/or "expedite"] the creation of metadata that's critical to leveraging digitized specimens for machine learning. The sections that follow cover background context, reviews the research design, reports results Next,  the  results  are presented, followed by a discussion. 
%items in square brackets [*xxx] are just other suggested words, e.g., select one or the other, or ....
%let's give a few data results, e.g., percentages, numbers..what will be emphasize in conclusion, key findings.
%abstract could say something about comparison to human results.

@@

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. The Darwin Core (DwC~\cite{biodiv_info_standards}, developed specifically to describe biological diversity data is among the most popular standard for such efforts. It is an extension of the Dublin Core's DCMI Metadata Terms \cite{dc_terms}. There is also the recent Audubon Core\cite{audub_core} supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens. These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is generally sparse and more prone to human error, placing limitations on a researcher seeking to apply machine learning to the image and the metadata for scientific research.

This limitation is magnified when trying get a handle on the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community advocated for data fitness standards \cite{chapman2020developing}. Even so, metadata properties used to record image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, and serves to justify a new initiative at Tulane University, where staff are manually capturing content for 22 metadata properties, to get a handle on the quality of digitized specimens drawn from the INHS collection and iDigBio. This current process is costly in terms of labor and time, and underscores the need to explore automatic metadata generation [*extraction].
